---
title: "Lab 1: An Introduction to R and RStudio"
author: " Armin Hatefi"
date: "Saturday, January 15, 2022"
output:
  pdf_document: default
  html_document: 
  theme: spacelab
toc: yes
urlcolor: BrickRed
---

***

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.align = "center")
```



This document shows you how to do simple tasks in RStudio.

# Starting RStudio

RStudio is a free software and is available on all the machines in most of the university computer labs. 
But this semester, we have to work on computer labs remotely. First, log in to your computer and to start RStudio.

**Note:** Throughout this lab we work with `RStudio Console` and type the following commands in Console. When we hit return key, the Console runs the command immediately. Later, we learn to type the commands in `RStudio Editor`  and then run the commands. 

# Help Commands

You can type either of 

```{r}
 help(log)
 ?log
 log(68)
```

To display the help file for the log (or any other) command. Type `help.start()` to start a help window. This is a way to list all the R commands and is **very useful**.

***

# Simple Tasks in R

RStudio is an interactive computing environment which you will use for data analysis. It can also be used as a calculator to perform simple tasks:

```{r}
5*3
sqrt(25)
4^2
abs(-7)
```

***

# Creating Variables

There are two assignment operators in R `<-` and `=`. They can be used interchangeably.

```{r}
x = 5
x
x <- 5
x
y = 3
z = 4
x + y + z
y
```

***

# Creating Vectors

```{r}
x = 4:12
x

y = seq(4,12,by=2)
y

y = seq(4,12,length=25)
y

height <- c(65,70, 66, 71, 66, 63)
height

height * 2.54


height[5]
height[c(2,5)]
height[-1]

length(height)

weight <- c(142,182,100,167,111,162)
weight

height/weight

name <- c('Marta',"John",'Doug','Sarah','Jen',"Jeff")
name

```

***

# Operations on Vectors

Create two vectors `x` and `y`, **having the same length** and see what happens when you do each of the following operations. For example, we use `height` and `weight` from above.

```{r}
x = c(5,2,1,4)
y = c(15,12,10,13)
x-y
y / x
x * y 
x^2
log(x)
cbind(x,y)
rbind(x,y)
```


***

# Making a Matrix and Having Access to it

```{r}
A = matrix(1:10,nrow = 5,ncol = 2)
A
B = matrix(1:10,nrow = 5,ncol = 2, byrow = T)
B
B[4,1]
B[1,2]
B[,1]
B[3,]
B[c(1,4),]
```

***

# Creating Data Frame

```{r}
students <- data.frame(name,height,weight)
students

students[4,2]
students[4,]
students[,2]

cars
cars[,1]
cars$speed
cars$dist
```

***


# Simple Random Sampling

In class, we learned how to take a simple random sample (SRS) 

- Table of Random Digits
- Applets on Internet 

We can use RStudio and take a simple random sample very easily. To take an SRS, we use command `sample`. 

Let's first load a built-in data set called `rivers`. 

```{r}
? rivers
sample(rivers,size=20)
sample(rivers,20)
sample(rivers,20,replace=T)
sample(x=rivers,size=20,replace=T)
```

Let's now make the problem a little more complicated. This time, we load `trees` data set.

```{r}
? trees
head(trees)
 dim(trees)
index <- sample(1:31,size=12)
index
trees[24,2]
trees12 <- trees[index,]
trees12
```

